---
title: "Mapping Geospatial objects in R "
output: html_notebook
---

## Advanced Geo Mapping in R using geospatial objects


Basic geospatial objects include:

* SpatialPolygons
* SpatialPoints
* SpatialLines
* SpatialPolygonsDataFrame
* SpatialPointsDataFrame
* SpatialLinesDataFrame


## Working with the `tigris` package
```{r}
if(!require(tigris)){install.packages("tigris", dependencies = TRUE); library(tigris)}
if(!require(sp)){install.packages("sp", dependencies = TRUE); library(sp)}
```

The tigris package lets you pull spatial data directly from the `US Census`. This data comes into R as a spatial object. To provide a basic overview of working with spatial object in R, we will use an example spatial object pulled with this package.

See the article [tigris: An R Package to Access with geodata on us census](https://journal.r-project.org/archive/accepted/walker.pdf)


### Using the `tracts` function in tigris package
```{r}
denver_tracts <- tracts(state = "CO", county = 31, cb = TRUE)
```
Using class on the object, we see that the function returned a SpatialPolygonsDataFrame object.
```{r}
class(denver_tracts)
```
These can be ploted
```{r}
plot(denver_tracts)
```

There are a number of additional methods for this specific class.

#### `bbox()` to display the bounding box of the spatial object
```{r}
bbox(denver_tracts)
```

With `is.projected` and `proj4string` we gather information about the currently applied coordinate system
```{r}
is.projected(denver_tracts)

proj4string(denver_tracts)
```

If a spatial object includes attribute data (i.e., is an object type *DataFrame), you can access a “slot” in the spatial object to pull out an attribute dataframe using @. For example, here’s the beginning of the dataframe for the denver_tracts spatial object:
```{r}
head(denver_tracts@data)
levels(factor(denver_tracts$AWATER))
```

We can overlay the fill attribute with different categorical variables:
```{r}
plot(denver_tracts, col = denver_tracts$AWATER)
```

### Plotting a spatial object using ggplot functions

There likely will also be cases when you have a spatial object but would like to plot it using ggplot-type mapping. In this case, you need to convert the spatial object to a dataframe, since ggplot will only input dataframes. You can use the fortify function from ggplot2 to do this. For example, to convert the Denver census tracts spatial object to a dataframe, you could call:

```{r}
# This function translates the sp objects into a grouped lat, long dataframe
denver_tracts_df <- fortify(denver_tracts)

# The compatible structure for a ggplot call was introduced
head(denver_tracts_df)

# We can plot this dataframe as usual using the geom_polygon layer
ggplot(denver_tracts_df, aes(lat, long, group = group)) +
  geom_polygon(fill = "lightblue", col = "black") +
  theme_void()
```

### Adapting and setting the coordinate system to work with

To go more in depth read trough [Mastering R from Roger Peng - Advanced Mapping](https://bookdown.org/rdpeng/RProgDA/mapping.html)

```{r}

```


