---
title: "R Notebook"
output: html_notebook
---

## Mapping Data in ggplot2

```{r}
require(ggplot2)
if(!require("maps")){install.packages("maps"); library(maps)}
us_map <- map_data("state")
```


## Drawing a map using a path geom
```{r}
require(dplyr)
us_map %>% 
  filter(region %in% c("north carolina", "south carolina")) %>% 
  ggplot(aes(long, lat)) + 
  geom_path()
```


## Drawing map using the geom_polygon()
```{r}
us_map %>% 
  filter( region %in% c("north carolina", "south carolina")) %>% 
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon( fill = "lightblue", col = "black")
```

## Applying a theme to seperate the picture from the data representation
```{r}
us_map %>% 
  filter(region %in% c("north carolina", "south carolina")) %>%  
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill ="lightblue", col = "black") + 
  theme_void()
```

## Plotting complete map data for the us regions
```{r}
# Plotting the map of all US States
baseplot <- us_map %>% 
  ggplot(aes(long, lat, group = group )) + 
  geom_polygon( fill = "lightblue", col = "black") + 
  theme_void()

# Highlighting north dako
map_nd <- us_map %>% 
  filter(region == "north dakota") %>% 
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon( fill = "red", col = "black")

```
### Mapping values onto the basemap

```{r}
library(dplyr)
library(viridis)

repub.votes <- votes.repub %>% 
  tbl_df() %>% 
  mutate(state = rownames(votes.repub),
         state = tolower(state)) %>% 
  right_join(us_map, by = c("state" = "region"))

```


```{r}
require(readr)
serial <- read_csv(paste0("https://raw.githubusercontent.com/",
                          "dgrtwo/serial-ggvis/master/input_data/",
                          "serial_podcast_data/serial_map_data.csv"))

# Checking the dataset
head(serial)

# Mutating the dataset to fit longitude, latitude format
serial <- serial %>% 
  mutate(long = -76.8854 + 0.00017022 * x, 
         lat = 39.23822 + 1.371014e-04 * y, 
         tower = Type == "cell-site")

serial %>% 
  slice(c(1:3, (n() - 3):(n())))

head(serial)
baltimore <- map_data('county', region = "maryland") %>% 
  filter(subregion %in% c("baltimore city", "baltimore"))

# Plotting the location map of baltimore as a subset of Maryland
maryland_plot <- map_data('county', region = "maryland") %>% 
  filter(subregion %in% c("baltimore city", "baltimore")) %>% 
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon(fill = "lightblue", col = "black") +
  theme_void()

maryland_plot

# Adding datapoints 
maryland_plot + 
  geom_point(data = serial, aes(group = NULL, color = tower)) + 
  scale_color_manual(name = "Cell tower", values = c("black", 'red'))

```

## Plotting with google Maps trough the `ggmap` package
```{r}
if(!require(ggmap)){install.packages("ggmap"); library("ggmap")}
require(ggplot2)
```

Basic Syntax for the map calls trough the API.

`get_maps()` to get a basic map for a location
```{r}
# This call returns a basic map object
Mosergasse <- get_map("Mosergasse 12/22, 1090 Wien")

# It can be plottet using the ggmap function
ggmap(Mosergasse)
```
```{r}
map_1 <- get_map("Estes Park", zoom = 12,
                 source = "google", maptype = "terrain")
ggmap(map_1) +
  theme_void() + 
  ggtitle("Estes Park Terrain View")
```

```{r}
moser_map <- get_map("Mosergasse 12/22, 1090 Wien", zoom = 15)
ggmap(moser_map) + 
  theme_void() + 
  ggtitle("Vienna Austria")
```

Adding information to the map
```{r}
require(dplyr)
require(ggplot2)

# Hybrid Map
map_1 <- get_map("Mosergasse 12, 1090 Wien",
                 zoom = 12, 
                 maptype = "hybrid",
                 source = "google") %>% 
  ggmap(extend = "device") + 
  theme_void()

# Terrain Map
map_2 <- get_map("Mosergasse 12, 1090 Wien",
                 zoom = 12, 
                 maptype = "terrain",
                 source = "google") %>% 
  ggmap(extend = "device") + 
  theme_void()

# Stamen Watercolor Map
map_3 <- get_map("Mosergasse 12, 1090 Wien",
                 zoom = 12, 
                 source = "stamen",
                 maptype = "watercolor") %>% 
  ggmap(extend = "device") + 
  theme_void()

library(gridExtra)
grid.arrange(map_1, map_2, map_3, nrow = 1)
```

Adding geoms to the map object
```{r}
get_map("Baltimore County", zoom = 10,
        source = "stamen", maptype = "toner") %>% 
  ggmap() + 
  geom_polygon(data = baltimore, aes(long, lat, group = group),
               color = "navy", fill = "lightblue", alpha = 0.2) + 
  geom_point(data = serial, aes(long, lat, color = tower)) + 
  theme_void()
  
```

```{r}
get_map("Goa, India", zoom = 10,
        source = "google", 
        maptype = "terrain") %>% 
  ggmap(extend = "device") + 
  theme_void()
```

```{r}
get_map("Baltimore County", zoom = 10,
        source = "google", maptype= "hybrid") %>% 
  ggmap()  + 
  geom_polygon(data = baltimore, aes(long, lat, group = group),
               color = "navy", fill = "lightblue", alpha = 0.4) + 
  geom_point(data = serial, aes(long, lat, color = tower)) + 
  theme_void() + 
  scale_color_manual(name = "Cell tower", values = c("black", "red"))
```
```{r}
get_map("Baltimore County", zoom = 10, 
        source = "stamen", 
        maptype = "toner") %>% 
  ggmap(extend = "device") + 
  geom_polygon(data = baltimore, aes(long, lat, group = group), col = "navy", fill = "lightblue", alpha = 0.2) + 
  geom_point(data = serial, aes(long, lat, color = tower)) + 
  theme_void() + 
  scale_color_manual(name = "Cell tower", values = c("black", "red"))

```
### Using the ggmap package to get geolocation using `geocode()` functionality

```{r}
geocode("One World Center")
```

```{r}
geocode("Supreme Court of the United States") %>% 
  get_map(zoom = 15) %>% 
  ggmap() +
  theme_void()
```
### Computing the distance given two points 

```{r}
mapdist("Baltimore, MD",
        "1 First St NE, Washington, DC") %>%
  select(from, to, miles)
```

Other functionalities can be found here [goole maps API Documentation] (https://developers.google.com/maps/documentation/distance-matrix/?csw=1)

