---
title: "Advanced DataVisualization with ggplot"
output: html_notebook
---

## Advanced DataVisualization with ggplot - Themes

Themes controll all non-data ink information for the plots.
These are all visual elements, not part of the data representation, and they consist of three parts:

- text
- line
- rectangle

They are accessed by a standard nameing convention in ggplot:

- element_text()
- element_line()
- element_rect()

We can not also change the visual attributes of these elements, but also their positions on the plot. 


## First we look at the visual elements
```{r}
plot <- ggplot(iris.wide, aes(Length, Width, col = Part)) + 
  geom_point() + 
  facet_grid(. ~ Species)

# This is the displayed plot in standard
plot
```

Now we can individually specify the elements we want to alter
```{r}
# We can completly remove a certain layer from the plot
plot

# "Text" Elements
plot + 
  theme(text = element_blank())
    
  # And alter the subsets of text elements on the plot
    # Title
    plot + theme(title = element_text(face = "italic", margin = 50, size = 24, colour = "blue"))
    
    # Axis legend title
    plot  + theme(legend.title = element_text(size = 17, 
                                              color = "darkgrey",
                                              face = "bold"))

    
# "Line" Elements
plot + theme(line = element_blank())
    
    # And their basic components `ticks`, `lines`, `grid`
    plot + theme(
                # "Red"
                axis.ticks = element_line(color = "red",
                                           size = 1, 
                                           linetype = 6, 
                                           lineend = "round") , 
                 # "Blue"
                 axis.line = element_line(color = "blue", 
                                          lineend = "round"), 
                 # 
                 panel.grid = element_line(linetype = 2,
                                           size = 1, 
                                           color = "red",
                                           lineend = "round")
    )

    
# "Rectangle" Elements, displaying the content
plot  + theme(rect = element_blank())

    # And their components `legend`, `panel`, `plot`
    
    # "Legend" Attributes
      # "Legend keys"
      plot + theme(legend.key = element_rect(
        fill = "lightgreen", 
        color = "red", 
        linetype = 3, 
        size = 0.3 
        
      ))
      
      # "Background"
      plot  + theme( legend.background = element_rect(fill = "red", 
                                                      color = "blue", 
                                                      size = 2.0))
      # "Position"
      # plot + theme(
      #   legend.position = element_rect()
      # )
      
      
  # "Plot" Attributes
  
      # "Background"
      plot + theme(
        plot.background = element_rect(fill = "lightgreen")
      )
      
      # 
      # plot + theme(
      #   plot.margin = element_rect()
      # )
  
  # "Panel" Attributes
      
      # "Plot Background"
      plot + theme(
        panel.background = element_rect(fill = "lightblue")
      )
      
      # "Border"
      plot + theme(
        panel.border = element_rect(color = "red", fill = "blue", size = 4)
      )
      
      # "stip.Background"
      plot + 
        theme(
        strip.background = element_rect(color = "blue", fill = "green", size = 2.5)
      )
    

```

## "Legend" - Positioning

1. We can set them with named attributes:
```{r}
# These are possible attributes to be set to position the legend

# Setting position by name
 
# "Bottom"
plot + theme( legend.position = "bottom" )

# "Left"
plot + theme( legend.position = "left" )

# "Top"
plot + theme( legend.position = "top" )


```

2. We can specify exact locations
```{r}
# Lower right corner, just above the x.axis labels
plot + theme( legend.position = c(0.92,0.15))

# upper left corner in the middle facet
plot + theme( legend.position = c(0.405, 0.87))
```

f