---
title: "R Notebook"
output: html_notebook
---

## Outline of the Websraper for the "IMDB Moviedatabase" 

Summary:


Packages applied:
- `rvest`



### Setting up the Data Frame "movie" to act as a temporary replacement for a Database
```{r}
require(dplyr)
movie <- data.frame(
          "ID" = character(),    #Will be set to the  ID string from the IMDb html adress
          "title" = character() ,
          "synomys_title" = character(), 
          "year" = numeric(),
          "dat_release" = as.Date(character()),
          "runtime_min" = numeric(),
          "directors" = character(), 
          "writer" = character(),
          "language" = factor(), 
          "box_opening" = numeric(), 
          "box_weekend" = numeric(), 
          "box_gross" = numeric(), 
          stringsAsFactors = FALSE)
```

```{r}
str(movie)
```

### Setting up a Date.Frame to hold the review information for a certain movie
```{r}
movie_review <- data.frame(
  ""
)
```

Movie related Information
    - Title of Movie
    - Release Date of Movie
    - Cast
      - Actor/ (Associated Figure)
    - Producer
    - Directors
    - Writers
    - Storyline
    - Synonyms(Title)
    - Language
    - BoxOffice_Opening
    - BoxOffice_Weekend
    - BoxOffice_Gross
  
  - Production Related
    - Runtime(min)
    - ProductionComp
    

  - Critics Reviews
    - Nominations
    - Metascore Critics Rating
    - Number of Critics Reviews
    - MPAA Rating
    
  - Public Reviews
    - Metascore Critics Rating
    - Number of Reviews(Public)
    - Number of Reviews(Critics)
    - Popularity Metrics
    - Summary Rating Statistics
      - Rated(male)
      - Rated(female)
      - Rated(< 18 & male)
      - Rated(18 & female)
      - Rated(18 - 29)
      - etc.... 

### CodeExample - Scraping Movie Rateings on IMDB
```{r}
# Scraping movie reviews from IMDB

# Collecting the source data
movie_html <- read_html("http://www.imdb.com/title/tt1490017/")
names(movie_html)
```
```{r}
# Scrape the movie review
avg_review <- movie_html %>% 
      html_node ("strong span") %>% 
      html_text() %>% 
      as.numeric()

# Scrape the cast
movie_cast <- movie_html %>%
  html_nodes("#titleCast .itemprop span") %>% 
  html_text()

# Get recent message board postings on this movie for 3 classes of review(loved, hated, bestvoted reviews)
review_best <- read_html("http://www.imdb.com/title/tt1490017/reviews?ref_=tt_urv")
review_loved <- read_html("http://www.imdb.com/title/tt1490017/reviews?filter=love")
review_hated <- read_html("http://www.imdb.com/title/tt1490017/reviews?filter=hate")

review_conc <-  list(review_best, review_loved, review_hated)

# For each review class gather the following information

# Parse out and then clean the rater names
rater_name <- review_best %>% 
  html_nodes("#tn15content div a") %>% 
  html_text()
rater_name <- rater_name[grep('.+', rater_name)]
  
# Parse out and then clean the rater_origin
rater_origin <- review_best %>% 
  html_nodes("small:nth-child(9)") %>% 
  html_text()

# Try advanced rater origin 
rater_origin_adv <- review_best %>% 
  html_nodes("small:nth-child(9) :nth-child(9) :nth-child(9) , small:nth-child(9), #tn15content div a") %>% 
  html_text()
rater_origin_adv[3]

rater_name <- rater_origin_adv[seq(2,length(rater_origin_adv), 3)]
rater_origin <- rater_origin_adv[seq(3,length(rater_origin_adv),3)]
data.frame(rater_name, rater_origin)
rater_origin





review_date <- review_best %>% 
  html_nodes("small:nth-child(10) , small:nth-child(11)") %>% 
  html_text()

review_date
review_class <- 

review_content <- review_best %>% 
  html_nodes("p:nth-child(46) , :nth-child(42), :nth-child(38), :nth-child(34), :nth-child(30), :nth-child(26), :nth-child(22), :nth-child(18), :nth-child(14), p:nth-child(10)") %>% 
  html_text()

review_content[grepl('[.+{10,}]+', review_content)]
review_title <- 
review_usefullnes <- 


```
### Collect the summary statistics on the movie Rating behaviour and clean it
```{r}
# Select the source and download html code
review_summary <- read_html("http://www.imdb.com/title/tt1490017/ratings?ref_=tt_ov_rt")

# Parse the data and select the relevant information
review_stats <- review_summary %>% 
  html_nodes("table:nth-child(11)") %>% 
  html_table(fill = TRUE)

# Parse the demographic breakdown for the rating behaviour
demog_stats <- review_summary %>% 
  html_nodes("table:nth-child(6) td") %>% 
  html_text()

dim (demog_stats) <- c(3)

# breakdown demog_stats
require(Hmisc)
rbind(demog_stats[1:3], demog_stats[4:6], demog_stats[7:9], demog_stats[10:12])
length(demog_stats)
split(, ceiling(seq_along(d)/20))
chunk2 <- function(x,n) split(x, cut(seq_along(x), n, labels = FALSE)) 
chunk2(demog_stas)
```

Create dataframe
```{r}
length(rater_name[grep(pattern = ".+", rater_name)])

data.frame(rater_name, rater_origin, review_date)
```


Summarize function
```{r}
print(avg_review)
print(movie_cast)
seq(1,10,2)