---
title: "R_GettingCleaning_WorkingExamples"
author: "Frank Fichtenmueller"
date: "2 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Short Example Workflows for Getting, Loading and Cleaning Datasets

This section introduces a short collection of simple exercises around Getting and Cleaning of Datasets. 

The Data in this part of the Repo focusses on:

- Code Base Files for specific Applications in **Getting and Cleaning of Datasets**
- Workexamples adressing individual Application Cases. They focus on real life application szenarios and the learnings derived from them. 



######Oscar Movie Database Example

This example highlights the process of joining.
```{r}
#Setting up a small dataset to work with

library(readr)

oscars <-"
name,movie,category
Adam McKay,The Big Short,Best Director
Alejandro González Iñárritu,The Revenant,Best Director
Lenny Abrahamson,Room,Best Director
Tom McCarthy,Spotlight,Best Director
George Miller,Mad Max: Fury Road,Best Director
Bryan Cranston,Trumbo,Best Actor
Matt Damon,The Martian,Best Actor
Michael Fassbender,Steve Jobs,Best Actor
Leonardo DiCaprio,The Revenant,Best Actor
Eddie Redmayne,The Danish Girl,Best Actor
Cate Blanchett,Carol,Best Actress
Brie Larson,Room,Best Actress
Jennifer Lawrence,Joy,Best Actress
Charlotte Rampling,45 Years,Best Actress
Saoirse Ronan,Brooklyn,Best Actress
"
oscars <- read_csv(oscars, trim_ws = TRUE, skip = 1)
oscars 

movies <-"
movie,length_mins
The Big Short,130
Star Wars: The Force Awakens,135
Brooklyn,111
Mad Max: Fury Road,120
Room,118
The Martian,144
The Revenant,156
Spotlight,128
"
movies <- read_csv(movies, trim_ws = TRUE, skip = 1)
movies 
```

Applying different join methods to compare the results
```{r}
# Comparing difference in passing datasets to `semi_join()`
semi_join(oscars, movies, by='movie')
semi_join(movies, oscars, by='movie')

# Combining the datasets and returning all rows from the oscars, and in both the oscars and movies column
left_join(oscars, movies, by='movie')

# Using the dplyr join functions, combine the columns from the oscars and movies data sets and return all 
# rows from the movies data set and all columns in both the oscars and movies columns.
right_join(oscars, movies, by='movie')

# 
inner_join(oscars, movies, by='movie')

# Using the dplyr join functions, return all rows from oscars data set where there are not matching values in 
# movies, only keeping the columns from the oscars data set.
anti_join(oscars, movies, by='movie')
```


###### NYCFlights Data
This example focusses on more complex summarisation tasks to answer still simple questions about the data.
```{r}
library(nycflights13); data(flights)
flights <- tbl_df(flights)

hist(flights$month)
summary(flights$month)
range(flights$month)
boxplot(flights$month)

hist(flights$day)
summary(flights$day)
range(flights$day)

longest_flights <- flights %>%
  select(month, day, carrier, air_time, distance) %>%
  filter(air_time > quantile(air_time, p=0.5, na.rm=TRUE)) %>%
  group_by(carrier) %>%
  summarize(n = count, mean_airtime = mean(air_time), dist_avg = mean(distance), dist_var = sd(distance)) %>%
  arrange(desc(mean_airtime), desc(dist_avg))

longest_flights



# Working on the mamals sleep datasets

library(readr)
library(dplyr)
library(ggplot2)

msleep %>% select(name:order)
msleep %>% select(starts_with('sl'))
msleep %>% select(ends_with('re'))
contains()
matches()
one_of()

# Select all columns from genus to conservation and filter the rows for mamals that sleep a total of more than 16 hours and have a body weight of greater than 1 kg
msleep
msleep %>% select(-(genus:conservation)) %>%
  filter(sleep_total > 16 & bodywt > 1)

# Create a new column called rem_proportion which is the ratio of rem sleep to total amount of sleep and create boxplots of the rem_proportion column split and 
# colored by the vore column. Create labels for the x and y axis.
msleep %>%
  mutate(rem_proportion = sleep_rem / sleep_total) %>%
  ggplot(aes(x = vore, y = rem_proportion, fill = vore)) + 
  geom_boxplot() + xlab("Type of diet") + 
  ylab("Proportion") + labs(title = "Proportion of rem sleep")

# Select all columns that start with the character string "sl" or ends with the character string "wt", create a new column called rem_proportion which 
# is the ratio of rem sleep to total amount of sleep, create a second column bodywt_grams which is the bodywt column in grams and filter for the rows 
# 20 to 30 in the msleep data set by numerical position.

Hint: Look at the slice() help file to filter for rows by numerical position.
msleep %>%
  select(starts_with("sl") , ends_with('wt')) %>%
  mutate(rem_proportion = sleep_rem / sleep_total, bodywt_grams = bodywt * 1000) %>%
  slice(10:30)


# To arrange (or re-order) rows by a particular column such as the taxonomic order, list the name of the column you want to arrange the rows by
msleep %>% 
  arrange(order)


# Select all columns names with the characters "sleep" and arrange the rows for the sleep_rem in a decreasing order.
# Hint: look at the ?arrange help file for the desc() option.
msleep %>%
  select(contains('sleep')) %>%
  arrange(desc(sleep_rem))


# Select three columns from msleep (name, order, sleep_total), arrange the rows in the sleep_total column in a descending order, and filter the rows 
# for mammals that sleep for a total of 16 or more hours.
msleep %>%
  select(name, order, sleep_total) %>%
  arrange(desc(sleep_total)) %>%
  filter(sleep_total >= 16)


# to compute the average number of hours of sleep, apply the mean() function to the column sleep_total and call the summary value avg_sleep.
msleep %>%
  summarise(avg_sleep = mean(sleep_total), length = n(), max_sleep = max(sleep_total), min_sleep = min(sleep_total), n_distinct = n_distinct(sleep_total))

# Summarize sleep_total column in the msleep data set with the average sleep, the minimum and maximum amount of sleep, and the total number of mammal
msleep %>%
  summarise(avg_sleep = mean(sleep_total), 
            min_sleep = min(sleep_total), 
            max_sleep = max(sleep_total), 
            n = n())

# Say we wanted to calculate the standard deviation of the body and brain weights for each of factor in the vore column
msleep$vore
msleep %>%
  ggplot(aes(x = vore, y = bodywt, fill = vore)) +
  geom_boxplot() + xlab("Type of Diet") + ylim(0,450)

# Split the msleep data frame by the taxonomic order, then for each taxonomic order summarize the sleep_total with the average sleep, 
# the minimum and maximum amount of sleep, and the total number of mammals in each order.

msleep %>%
  group_by(order) %>%
  summarize(avg_sleep = mean(sleep_total),
            min_sleep = min(sleep_total),
            max_sleep = max(sleep_total),
            total_n = n())

# Joining Dataframes with dplyr
x <- 

left_join(x,y, by='x1') # Join matching rows from y -> x
right_join(x,y, by='x1') # matching from x -> y
inner_join(x,y, by=) # Join data. retain only complete data
full_join(x,y, by=) # Join data, keep everything

# Filtering joins
semi_join(x,y, by=) # all rows in x have to match y
anti_join(x,y, by=) # all rows in x have to not-match y


#

###Advanced Work with dplyr, ggplot, tidyr on the nycflights13 dataset

First loading the required libraries and the datasets

######
library(dplyr); library(nycflights13); library(ggplot2); library(tidyr)
