---
title: "lubridate - Basic Syntax"
output: html_notebook
---


```{r}
if(!require(lubridate)){install.packages("lubridate", dependencies = TRUE)}
```

## Basic Syntax for the Lubridate Package in R

### Parsing Date from multiple sourcetypes
Given a string of different timerepresentations, we will explore all possibilities in the package to extract them into POSIX Datetime objects
```{r}
time_str <- list("1997/12/10", "01/01/1985", "21st January 1887", "01-31-1900", "March 12 87", "19900112")

# We can transform them, giving the order of year, month, day
ymd(time_str[1])
dmy(time_str[2])
dmy(time_str[3])
mdy(time_str[4])
mdy(time_str[5])
ydm(time_str[6])
```

### Parsing DateTime Objects:
```{r}
dtime_str <- list("12011900:18:00:34", "12/01/86:03:45 am","12-1-86:03:45am") 

# converting both Date and Time information from the list
mdy_hms(dtime_str)
mdy_hm(dtime_str[2])
mdy_hm(dtime_str[3])
```

### Using the `pase_date_time()` function

parse_date_time parses an input vector into POSIXct date-time object. It differs from strptime
in two respects. First, it allows specification of the order in which the formats occur without the
need to include separators and "%" prefix. Such a formating argument is refered to as "order". Second,
it allows the user to specify several format-orders to handle heterogeneous dat-time character represantions.

```{r}
# First get the information on the date-time representations of your data
str(time_str)
```

When working with such a lot of many different representations in a single dataset, its easier to pass them all as attributes.
```{r}
# Set the char representations 
orders <- c("ymd", "mdy", "dmy", "dmy", "mdy", "ydm")


# Call the parse_date_time() with the orders attributes
parse_date_time(time_str, orders = orders)
```

The Timezone can be automatically set, given the indivual customized standard of your system. Or you assign them.
```{r}
# Setting a new standard timezone to be applied to all DateTime Objects
parse_date_time(time_str, orders, tz="Europe/Vienna")
parse_date_time(time_str, orders, tz="Australia/Brisbane")

# Setting individual timezones for all entries
time_str_Pos <- parse_date_time(time_str, orders)

## First collect them in a list
time_str.tzs <- c("Europe/Vienna", "USA/New York", "Australia/Sydney", "Asia/New Delhi", "Europe/Moskow", "Asia/Peking")
## Then assign them onto the timeobject


# Set a Timezone as current TZ
Sys.setenv(tz = "Europe/Vienna")
tz(now())
now()
```

Quickly transform a timezone object into a different timezone

```{r}
tstamp <- now()
with_tz(tstamp, "Europe/London")
with_tz(tstamp, "Asia/Tokyo")
```


### Transforming multiple Timezone DateTime Objects into a localzone representation
```{r}
require(lubridate)
require(dplyr)

df = data.frame(timestring = c("2015-12-12 13:34:56", "2015-12-14 16:23:32"),
                localzone = c("America/Los_Angeles", "America/New_York"), stringsAsFactors = F)

df$moment = as.POSIXct(df$timestring, format="%Y-%m-%d %H:%M:%S", tz="UTC")

df = df %>% rowwise() %>% mutate(localtime = force_tz(moment, localzone))

df
```


### Importing DateTime Objects from multiple Timezones into a coherent representation

```{r}
# Given this information
dtime_Oakland <- list("12/04/2016 4:35pm")
dtime_NewYork <- list("04-03-16 12:45am")
dtime_Munich <- list("01.09.2016 16:45")

# First we transform the lists into DateTime object
dtime_Oakland <- mdy_hm(dtime_Oakland)
# We then have to add the local timezone information
tz(dtime_Oakland) <- "Pacific/Auckland"
tz(dtime_NewYork) <- 
dtime_Oakland
Sys.timezone()
now()

```


### Calculating Periods
```{r}
# Getting a current Timestamp
t_stamp <- now()

# Adding a given Timeperiod
next_stamp <- t_stamp + hours(3)

# Calculating time Difference
time_diff <- next_stamp - t_stamp
time_diff
```


This creates an object of class `difftime`, that has special attributes associated to allow advanced Calulations on TimeDifferences
```{r}
class(time_diff)
attributes(time_diff)
```
We can directly alter
```{r}
time_diff
```

#### Creating Periods 

period creates a period object with the specified values. period provides the behaviour of period
in a way that is more suitable for automating within a function

```{r}
# Creating a period given the num of units, and the units to be used
period(360, units="minutes")

# You can pass the following argument as unigs
units <- c('seconds', 'minutes', 'hours', 'days', 'weeks', 'months', 'years')

for(unit in  units){
  print(sprintf("Using unit '%s' we get a period of value of %s :", unit, period(30, unit)))
}
```

Within a Period object, time units do not have a fixed length (except for seconds) until they are
added to a date-time. The length of each time unit will depend on the date-time to which it is added.
For example, a year that begins on 2009-01-01 will be 365 days long. 
A year that begins on 2012-
01-01 will be 366 days long. When math is performed with a period object, each unit is applied
separately. How the length of a period is distributed among its units is non-trivial. For example,
when leap seconds occur 1 minute is longer than 60 seconds.
Periods track the change in the "clock time" between two date-times. They are measured in common
time related units: years, months, day

```{r}
t_timestamp <- now()
p_30 <- period(30)

# Adding different periods to the current time
values <-  c("second", "minute", "hour", "day", "week", "month", "year")

for(value in values){
  print(sprintf("Adding 30 %s to the current time gives us: %s", value, (t_timestamp + period(30, units=value))))
}
```

Periods can be assigned without value. Then they get their value from the timestamp they are assigned to
```{r}
timestamps <- list(second(now()), minute(now()), hour(now()), day(now()), year(now()))
for(unit in timestamps){
  print(unit + p_30)
}
```

### Setting different periods
```{r}
period(0, 'seconds')
period(secons=12, minutes=5, hour=10)
period(c(1,-60), c('hour', 'minute'))
now() + period(c(1,-60), c('hour', 'minute')) 

# Using shorthand notation
period("2d 2H 2M 2S")
period("2mins 4days 6hours")
period("2min 4d 24h")

# Blank values default to 1
period("sec min day")

# Repeated units get added up
period("day day minute 10min")
```
### Setting quick `durations()` to manipulate date-time objects
```{r}
t_stamp <- now()

# Move the time forward by a given amount
t_stamp + dseconds(40)
t_stamp + dminutes(65)
t_stamp + dhours(2.5)
t_stamp + ddays(0.3)
t_stamp + dweeks(2)
t_stamp + dyears(1.2)

## Using very small durations with
m_stamp <- dmicroseconds(10)
m_stamp + dmicroseconds(100)
m_stamp + dmilliseconds(10)
m_stamp + dnanoseconds(1)
m_stamp + dpicoseconds(1)

# All duration objects can be added or subtracted
now() + ddays(10) - dhours(5)
```

### Calling quick period objects for quick datamanipulation

```{r}
days(5)
class(days(5))

# Available period objects for all types
days(1)
minutes(10)
hours(4)
seconds(10)
weeks(2)
months(5)
years(2)
#etc...
```

### Working with the `period class` in lubridate

Period is an S4 class that extends the Timespan-class class. Periods track the change in the "clock
time" between two date-times. They are measured in common time related units: years, months,
days, hours, minutes, and seconds. Each unit except for seconds must be expressed in integer values.

```{r}

```


The exact length of a period is not defined until the period is placed at a specific moment of time.
This is because the precise length of one year, month, day, etc. can change depending on when it
occurs due to daylight savings, leap years, and other conventions. A period can be associated with
a specific moment in time by coercing it to an Interval-class object with as.interval or by
adding it to a date-time with "+".
Periods provide a method for measuring generalized timespans when we wish to model clock times.
Periods will attain intuitive results at this task even when leap years, leap seconds, gregorian days,
daylight savings changes, and other events happen during the period. See Duration-class for an
alternative way to measure timespans that allows precise comparisons between timespans


### Working with specific time periods

```{r}
# Get the quarter dates for a given timestamp
quarter(now(), with_year = TRUE)

# Get the semester
semester(now(), with_year = TRUE)
```

When measuring the duration of a given period we can use the `interval()` function from lubridate
```{r}
# First create a interval object
int <- interval(now(), dmy("11/09/1982"))

# Then get the duration values from it
time_length(int,unit = "years")
```



### Additional transformations on DateTime Objects
```{r}
# Rounding timestamps
round_date(now(), 'minutes')
round_date(now(), 'hours')
round_date(now(), 'weeks')
round_date(now(), 'day')

# Rounding down using floor_date
floor_date(now(), 'hours')
floor_date(now(), 'days')

# Rounding up using ceiling_date
ceiling_date(now(), 'hours')
ceiling_date(now(), 'weeks')

```

#### Unconventional period values include:

- `bimonth`
- `quarter`
- `halyear`
- `year`


### Assigning TimeStams

These functions enable a quick and human friendly annotation with timeinformation.
```{r}
D <- ymd("2010-04-05") - days(1:5)
stamp("March 1, 1999")(D)
sf <- stamp("Created on Sunday, Jan 1, 1999 3:34 pm")
sf(D)
stamp("Jan 01")(D)
stamp("Sunday, May 1, 2000", locale = "en_US")(D)
stamp("Sun Aug 5")(D) #=> "Sun Aug 04" "Sat Aug 04" "Fri Aug 04" "Thu Aug 04" "Wed Aug 03"
stamp("12/31/99")(D) #=> "06/09/11"
stamp("Sunday, May 1, 2000 22:10", locale = "en_US")(D)
stamp("2013-01-01T06:00:00Z")(D)
stamp("2013-01-01T00:00:00-06")(D)
stamp("2013-01-01T00:00:00-08:00")(force_tz(D, "America/Chicago"))

```

