---
title: "R_Programming_SequencesOfNumbers"
author: "Frank Fichtenmueller"
date: "1 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Working with sequences of Numbers in R

Different approaches to generating basic sequences
```{r}
1:10
pi:10
seq(1:10)
seq(5,10,0.5)
seq_30 <- seq(5,20, length=30)
length(seq_30)

# Enumerating a given vector e.g. for use in iterations
seq_along(seq_30)
```

Creating replications of a given sequence
```{r}
# Creating a zero vector with length x=20
rep(0, 20)

# Repeating a certain sub vector a given times
rep(c(1,2,3), times=10)

# Repeating a sequence of items a given times each
rep(c(1,2,3), each=5)
```

Indexing with boolean arrays
```{r}

```

Returning the indices of a selection

We can use the `which()` function to return an object representation of a subsets index location.
```{r}
# Generating an example matrix
m <- matrix(1:60, ncol = 6)

# indexing out all values above 24
m[m > 24]

# As this returns a vector representation, we might not like it
which(m>24)
```

Using the `any()` function to return a logical evaluation if any of the items evaluate to TRUE

```{r}
any(m <0)
```

Using the `all()` funt

