---
title: "R_TextProcessing_RegEx"
author: "Frank Fichtenmueller"
date: "1 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## TextProcessing and Regular Expressions in R


#####Scope of this document:
Describe specific functionalities and their packages for the topics of `Textprocessing` and the application of `Regular Expressions` to string selection. 

*** 

#####This file contains sections on the following topics:
* Basic Text processing functions
* Basic Text Manipulation Functions
* Regular Expressions in R
* the `RegEx` Functions
* the `stringr` package

### Basic commands for Text processing
Text is represented as `string` objects. These can be split, recombined, changed and altered by all algorithmic functions. 

```{r}
# Concatening Strings
paste("Square", "Circle", "Triangle")

# sep defines the string to use as binding element
paste("Square", "Circle", "Triangle", sep="+")

# Shortcut to combining without any space
paste0("Square", "Circle", "Triange")
```

This also works when passing a string to paste()
```{r}
# When pasting a vector, each element is pasted after the other
s <- c("Square", "Circle", "Triangle")
paste("my favorites is a", s)

# The string can be injected anywhere
city <- c("Tokyo", "London")
times <- c('best', 'worst')
paste("The", times," city was", city)
```

specify the `collapse=" "` argument to `paste()` to collapse into a single string.
Enables to print all at the same time.
```{r}
paste0(c("Square", "Circle", "Triangle"), collapse=" ")
```

#### Functions to alter the text

* `tolower()`
* `toupper()`
* etc....

(!!!! TODO: ENTER ALL RELATED FUNCTIONS HERE)


### Regular Expressions
A regular expression is a string that defines a pattern that could be contained within another string. A regular expression can be used for searching for a string, searching within a string, or replacing one part of a string with another string. In this section I might refer to a regular expression as a regex, just know that they're the same thing.

Regular expressions use characters to define patterns of other characters.


#### Applying a regExp with the `grepl()` function in the Base Package

```{r}
regular_expression <- "a"
regular_expression2 <- "u"
string_to_search <- "Maryland"

# Searching 
grepl(regular_expression, string_to_search)
grepl(regular_expression2, string_to_search)
```

Like this we can easily search trough strings to identify matching patterns. To this end we use a list of meta characters,
to match on individual patterns, and exclude all unimportant patterns from our search.

These include:

* `.` - Any character other than a `\n`
* `+` - **One or more** of the preceeding character are present
* `*` - **Zero or more** of the preceeding character are present
* `{}` - **Exact** numbers of times preceeding character should be present
* `{2,5}` - **Range of times** preceeding characer should be present
* `




```{r}
# Example dataset
head(state.name)

# Applying a couple of different string matching criteria
grepl(".", "Maryland")

```

