---
title: "R_Working with Regular Expressions in R"
author: "Frank Fichtenmueller"
date: "1 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## TextProcessing and Regular Expressions in R

External Ressources to add:
[Tidy Text Mining with R](http://tidytextmining.com/)

#####Scope of this document:
Describe specific functionalities and their packages for the topics of `Textprocessing` and the application of `Regular Expressions` to string selection. 

*** 

#####This file contains sections on the following topics:
* Basic Text processing functions
* Basic Text Manipulation Functions
* Regular Expressions in R
* the `RegEx` Functions
* the `stringr` package
* the [tidytext package](https://github.com/juliasilge/tidytext)

### Basic commands for Text processing
Text is represented as `string` objects. These can be split, recombined, changed and altered by all algorithmic functions. 

```{r}
# Concatening Strings
paste("Square", "Circle", "Triangle")

# sep defines the string to use as binding element
paste("Square", "Circle", "Triangle", sep="+")

# Shortcut to combining without any space
paste0("Square", "Circle", "Triange")
```

This also works when passing a string to paste()
```{r}
# When pasting a vector, each element is pasted after the other
s <- c("Square", "Circle", "Triangle")
paste("my favorites is a", s)

# The string can be injected anywhere
city <- c("Tokyo", "London")
times <- c('best', 'worst')
paste("The", times," city was", city)
```

specify the `collapse=" "` argument to `paste()` to collapse into a single string.
Enables to print all at the same time.
```{r}
paste0(c("Square", "Circle", "Triangle"), collapse=" ")
```

#### Functions to alter the text

* `tolower()`
* `toupper()`
* etc....

(!!!! TODO: ENTER ALL RELATED FUNCTIONS HERE)


### Regular Expressions
A regular expression is a string that defines a pattern that could be contained within another string. A regular expression can be used for searching for a string, searching within a string, or replacing one part of a string with another string. In this section I might refer to a regular expression as a regex, just know that they're the same thing.

Regular expressions use characters to define patterns of other characters.


#### Applying a regExp with the `grepl()` function in the Base Package

```{r}
regular_expression <- "a"
regular_expression2 <- "u"
string_to_search <- "Maryland"

# Searching 
grepl(regular_expression, string_to_search)
grepl(regular_expression2, string_to_search)
```

Like this we can easily search trough strings to identify matching patterns. To this end we use a list of meta characters,
to match on individual patterns, and exclude all unimportant patterns from our search.

These include:

* `.` - Any character other than a `\n`
* `+` - **One or more** of the preceeding character are present
* `*` - **Zero or more** of the preceeding character are present
* `[a-z]` - **range** of characters
* `[1-9]` - **range** of integers
* `[xyz]` - **set** of characters
* `[^xyz]` - **negation** of set
* `\n` - **newline**
* `{}` - **Exact** numbers of previous
* `{2,5}` - **Range** of numbers of previous 
* `{2,}` - **More than** 2 of previous
* `^` - matches the **beginning** of a word
* `$` - matches **End** of a word
* `|` - matches **Either the first OR the second regEx**


**Introducing a final usecase** 
```{r}
# Matching all statenames in the US that begin, or end in a vowel
start_end_vowel <- "^[AEIOU]{1}.+[aeiou]{1}$"
vowel_state_lgl <- grepl(start_end_vowel, state.name)
state.name[vowel_state_lgl]
```

```{r}
# Matching Beginning or End of word
grepl("^[ab]+$", c("bab", "aab", "abc"))

# Matching OR
grepl("North|South", c("South Dakota", "North Carolina", "West Virginia"))


```



There are additionall `character sets` that come built into R:

* `\\w` - words specify any letter, digit or _
* `\\d` - digits specify any number(0-9)
* `\\s` - whitespace character specify line breaks, tabs, spaces

These have their compliments in capitalized form:
* `\\W` - not words
* `\\D` - not digits
* `\\S` - not whitespace

Lets look at some examples:
```{r}
grepl("\\w", "abcdefghijklmnopqrstuvwxyz0123456789")


```

##### Creating unique character sets

* Use `[]` to define a unique character set to match
* Use `^` to get the compliment of the statement following
```{r}
grepl("[aeiou]", "rhythms")

grepl("[^aeiou]", "rhythms")

grepl("[a-i]", "rhythms")

grepl("[1-5]", "welcome2014-party")

```

##### Matching punctuation characters


Putting two backslashes before a punctuation mark that is also a metacharacter indicates that you are looking for the symbol and not the metacharacter meaning


```{r}
grepl("\\+", "tragedy + time = humor")
```


```{r}
# Example dataset
head(state.name)

# Applying a couple of different string matching criteria
grepl(".", "Maryland")

```

#### Other basic regEx functions in R

* `grepl()` - grep logical function
* `sub(regex,num_found, string)` - replace the (num_found) instance of regex found in string
* `gsub(regex, string)` - replace all instances of regex in string
* `strsplit()` - split up strings according to provided regex

```{r}
# grepl()
grepl("[Ii]", c("Hawaii", "Illinois", "Kentucky"))

# grep()
grep("[Ii]", c("Hawaii", "Illinois", "Kentucky"))

# sub(
sub("[Ii]", "1", c("Hawaii", "Illinois", "Kentucky"))

# gsub()
gsub("[Ii]", "1", c("Hawaii", "Illinois", "Kentucky"))

# strsplit()
two_s <- state.name[grep("ss", state.name)]
two_s
strsplit(two_s, "ss")

```

### Basic Text manipulation functions






### The `stringr package`

Its a part of the Tidyverse group of packages from Hadley Wickham. It provides a "data first" approach to functions involving regex.

Usually the function calls are of sort `(string, regex)`. This is to facilitate the use with the `magrittr %>%` functionality, which passes the data as the first argument to following functioncalls. 
And most functions in stringr start with **"str_"**

There are the following functions:

* `str_extract()` - returns the sub string of a string that matches the regex
* `str_order()` - returns a numeric vector that corresponds to the alphabetical order of the strings in the provided vector
* `str_pad()` - pads strings with other characters. useful when to be eventually printed for a person to read.
* `str_to_title(cases)` - prints strings with first letter capitalized
* `str_trim()`- deletes whitespace from both sides of a string
* `str_wrap()` - inserts newlines in strings so that when the string is printed each line's length is limited
* `word()` - allows you to index each word in a string as if it were a vector.

Some examples for the function calls:
```{r}
require(stringr)

# Creating a state table to work with
state_tbl <- paste(state.name, state.area, state.abb)
head(state_tbl)

# Extracting all state area codes from the table 
str_extract(state_tbl, "[0-9]+")

# Alphabetical ordering of state.names
str_order(state.name)

# Padding a string
str_pad("Thai", width = 8, side = "left", pad = "-")

# Trimming whitespace
to_trim <- c("   space", "the    ", "    final frontier  ")
str_trim(to_trim)

# Wrapping text
pasted_states <- paste(state.name[1:20], collapse = " ")
cat(str_wrap(pasted_states, width = 80))

# indexing into a string like in a vector
a_tale <- "It was the best of times it was the worst of times it was the age of wisdom it was the age of foolishness"

word(a_tale, 2)
word(a_tale, end = 3)
word(a_tale, start = 11, end =15)

```

