---
title: "Example - The NYC Flights Dataset"
author: "Frank Fichtenmueller"
date: "30 October 2016"
output: html_document
---
Here i am going trough a sample analysis of the content of the NycFlights13 Dataset
```{r, include=FALSE}
#Loading the library and the dataset in dplyr df_table format
#if(!require("nycflight13") == TRUE) {install.packages("nycflights13"); library(nycflights13)}
library(nycflights13); library(ggplot2); library(tidyr)
if(!require("dplyr") == TRUE) {print('required packages not installed, will install now');install.packages("dplyr")} else {print('package installed and loaded')}
flights <- tbl_df(flights)
```
First we look get an overview of the dataset
 * Looking at the dimension of the table
 * The variables included (name, type, factor levels if categorial)
 * The timerange of the dataset
 * The summary statistics for each variable
 * The amount of missing data
 * 

### Inspecting the Question of Flight Delay
After subsetting the dataset on dep_delay values of below 240 to cut outliers,
```{r, include=FALSE}
######Filtering out large delays(outliers)
flights <- flights %>% filter(dep_delay < 240)

```
We can take a look at a quick summary of flight delay related measures.  
(Departure Delay and Arrival Delay)
```{r densityFunction, fig.width=8, fig.height=6, error=FALSE}
###Summarise each
flights %>%
  select(dep_delay, arr_delay) %>%
  summarise_each(funs(mean))

### To remove NAs from the calculation we can add the na.rm attribute. Here we pass the . to transport all current variables into the function
flights %>%
  select(dep_delay, arr_delay) %>%
  summarise_each(funs(mean(.,na.rm=TRUE,)), matches('delay'))


# Scaleing the data with tidyr
flights %>% ggplot(aes(x=dep_delay)) + geom_density()
flights %>% ggplot(aes(x=arr_delay)) + geom_density()
```  

Now, we can start looking at the density attributes of the variables to get a first understanding about possible correlations in the dataset.  

```{r, hightlight=TRUE}

# In order to plot multiple variables onto the same plot it is good practice to first gather() them into factor levels of one individual variable. 
# This can then be color coded onto the graph

flights %>% gather(delay, value, dep_delay, arr_delay) %>%
  ggplot(aes(x=value, fill=delay)) + geom_density(alpha=.3)


# This can also be applied with tidyr to represent the data in nice interpretational tables. 
flights %>%
  gather(measure, value, dep_delay, arr_delay, air_time, distance) %>%
  select(value, measure) %>%
  filter(!is.na(value)) %>%
  group_by(measure) %>%
  summarise_each(funs(mean, median, min, max, sd))

# Using this function to create a function to iterate over a list of variable names in the datapool. 
for (var in c("dep_delay", "arr_delay")) {
  flights %>%
    ggplot(aes_string(x=var)) +
    geom_histogram()
}

```

### Destinations from NYC
Next lets look at the destinations from NYC. 
In order to make proper sense of this data, we need to add the geolocation data for all the airports in the dataset. 
This information can be found from the [openflights project](www.github.com/jpatokal/openflights). We will download and import the data, select
the information we need and attach it into a combined dataset to work with. 

Just looking at the given data so far, there is limited information available, about the 
```{r}
flights %>%
  select(carrier, origin, dest, distance) %>%
  group_by(dest) %>%
  summarise(num_flights = n(), avg_distance = mean(distance)) %>%
  arrange(desc(num_flights)) %>%
  slice(1:40) %>%
  ggplot(aes(x=dest, y=num_flights)) + geom_()
  
```


