---
title: "WebScraper for Google scholar with rvest"
output: html_notebook
---

## Example Application of rvest scraping to Scholar Search on Google scholar

Summary:


Scope of the scraping:
Given the name of a person, the system shall return:

- Research Affiliation of the Person
- Articles written by the person
- Number of Citations
- Coauthors to the articles
- Affiliation of coauthors
- Number of papers published (Grouped by research area)
- 

Packages applied:

- rvest

#### Loading required libraries
```{r}
require(rvest)
require(ggplot2)
```

#Return google scholar page for the Person searched
```{r}
source_html <- read_html("https://scholar.google.com/citations?user=sTR9SIQAAAAJ&hl=en&oi=ao")
source_html_enh <- read_html("https://scholar.google.com/citations?user=sTR9SIQAAAAJ&hl=en&oi=ao")

```


# How many citations does the person have
```{r}
citation <- source_html %>% 
  html_nodes("gsc_a_b .gsc_a_c") %>% 
  html_text() %>% as.numeric()

citation
```

```{r}
citation <- source_html %>% 
  html_nodes(".gsc_a_ac") %>% 
  html_text() %>% as.numeric()

articles <- source_html %>% 
  html_nodes(".gsc_a_at") %>% 
  html_text()

coauthors <- source_html %>% 
  html_nodes(".gs_gray:nth-child(2)") %>% 
  html_text()

# Second implementation
  page <- read_html("https://scholar.google.com/citations?view_op=list_colleagues&hl=en&user=sTR9SIQAAAAJ")
  Coauthors = page%>% html_nodes(css=".gsc_1usr_name a") %>% html_text()
  Coauthors = as.data.frame(Coauthors)
  names(Coauthors)='Coauthors'

journal <- source_html %>% 
  html_nodes(".gs_gray:nth-child(3)") %>% 
  html_text()

year <- source_html %>% 
  html_nodes("#gsc_a_b .gsc_a_h") %>% 
  html_text() %>% as.numeric()

```

```{r}
# Creating the dataframe 
df <- data.frame("article" = articles, 
                 "coauthors" = coauthors,
                 "journal" = journal, 
                 "citations" = citation,
                 "year" = year,
                 stringsAsFactors = FALSE)
str(df)
head(df, 40)
```


#### Create Visualizations for dashboard context
```{r}
barplot(df$citations, main="Citations per Paper", ylab="Number of citations", col="blue", xlab="")
```
### Total Citations by year
```{r}
df %>% 
  group_by(year) %>% 
  summarise(total_citation = sum(citations)) %>% 
  plot()
```

