---
title: "Shiny with HTML and CSS"
output: html_notebook
---
```{r}
require(shiny)
```

## Using the `tags` to write the input to the shiny app

```{r}
# fluidPage writes an initial #body div to the page
ui <- fluidPage(
        
        # We can use nested #row divs to nest the content we want to put in a subset
        fluidRow(
                # Use the `tags$` to select the tags we want to use
                h2("This is an h2 header"),
                p("This text is displayed as a paragraph, followed with a ", a("link", href="http://google.com"))
        ),
        
        # Now we add a second row element to display
        fluidRow(
                h2("The second Header in a new row"),
                p("This is my new row.", br(),"it contains a break, and another", a("link", href="google.com"))
        ),
        
        # Creating the footer
        fluidRow(
                tags$footer(
                        h2("This is the footer")
                )
        )
)

server <- function(input, output){
        
}

shinyApp(ui, server)

```

```{r}
f1 <- function(inside, out){print("1")}
f2 <- function(inside, out){print("2")}
f3 <- function(inside, out){print("3")}
f4 <- function(inside, out){print("4")}
f5 <- function(inside, out){print("5")}

# Creating a list of functions
functions <- list(f1, f2, f3, f4)

# Calling the functions
functions[[1]]()
functions[[2]]()
```
### Creating a layout

placing all elements passed on x and y dimensions.
We can stack element using the z dimension. 

These fucntion are just there to space the elements in a layout. These are built on top of 
googles `bootstrap` html library. 

PERFECT!!!!

- `fluidRow`    : Adds rows to the body of the html
- `column`      : Adds the elements in a column as a subset of the 


The `column` takes the same arguments as in bootstrap:

- the width and the offset for each column out of 12

```{r}
# An example code here:
ui <- fluidPage(
        
        # Specify a header row
        titlePanel("SMART Recruiting"),
        
        # Set the header row
        fluidRow(
                # Creat the logo div
                tags$div(class = "logo", "Logo placeholder"),
                
                # Create the subtitle div
                tags$div(class = "subtitle", "Finding the right people now")
                ),
        
        
        # Specify a row to be separated into three columns 
        fluidRow(
                column(3,
                       # This part describes the content of the column
                       h4("This is the column header"),
                       p("This text goes in the column body")
                       
                       ),
                column(6,
                       # This part describes the content of the column
                       h4("This is the column header"),
                       p("This text goes in the column body")
                       ),
                column(3,
                       # This part describes the content of the column
                       h4("This is the column header"),
                       p("This text goes in the column body")
                       )
                ),
        
        # Set a column element to be offset all the way to the right
        fluidRow(
                column(4, offset = 8)
        )
)

server <- function(input, output){}

shinyApp(ui, server)
```


Working with the `navbar page` layout in Shiny

- use `navBar
- use the `navbarMenu` to create dropdown menue
```{r}
ui <- navbarPage(title="SMART Recruiting", 
                 
                 # First page
                 tabPanel("tab1", 
                          fluidRow(
                                  column(3,
                                         # Content of column 1
                                         tags$article(
                                                h2("header column 1"),
                                                p("The content for the article 1")
                                         )),
                                  column(3,
                                         tags$article(
                                                h2("header column 1"),
                                                p("The content for the article 1")
                                                )),
                                  column(3,
                                         tags$article(
                                                h2("header column 1"),
                                                p("The content for the article 1")
                                                )),
                                  column(3,
                                         tags$article(
                                                h2("header column 1"),
                                                p("The content for the article 1")
                                                ))
                          )), 
                 
                 # Second page
                 navbarMenu(title = "tab2", 
                          tabPanel("Subtab1", 
                                   "content"),
                          
                          tabPanel("Subtab2",
                                   "content"),
                          tabPanel("Subtab3",
                                   "content")
                          ), 
                 
                 # Third page
                 tabPanel("tab3", 
                          "contents"
                          )
                   
                   )

server <- function(input, output){}

shinyApp(ui, server)
```




## Working with the `dashboardPage()` layout:

This layout comes with a extra package, available from the package `shinydashboard`
```{r}
require(shiny)
require(shinydashboard)


ui <- dashboardPage(skin = "yellow",
        
        # Create header
        dashboardHeader(title="SMART Recruiting",
                        dropdownMenu(type = "messages",
                                          messageItem(
                                            from = "Sales Dept",
                                            message = "Sales are steady this month."
                                          ),
                                          messageItem(
                                            from = "New User",
                                            message = "How do I register?",
                                            icon = icon("question"),
                                            time = "13:45"
                                          ),
                                          messageItem(
                                            from = "Support",
                                            message = "The new server is ready.",
                                            icon = icon("life-ring"),
                                            time = "2014-12-01"
                                          )
                                        ),
                        dropdownMenu(type = "notifications",
                                          notificationItem(
                                            text = "5 new users today",
                                            icon("users")
                                          ),
                                          notificationItem(
                                            text = "12 items delivered",
                                            icon("truck"),
                                            status = "success"
                                          ),
                                          notificationItem(
                                            text = "Server load at 86%",
                                            icon = icon("exclamation-triangle"),
                                            status = "warning"
                                          )
                                        ),
                        dropdownMenu(type = "tasks", badgeStatus = "success",
                                          taskItem(value = 90, color = "green",
                                            "Documentation"
                                          ),
                                          taskItem(value = 17, color = "aqua",
                                            "Project X"
                                          ),
                                          taskItem(value = 75, color = "yellow",
                                            "Server deployment"
                                          ),
                                          taskItem(value = 80, color = "red",
                                            "Overall project"
                                          )
                                        )
                        
                        ),
        
        # Create Sidebar
        sidebar <- dashboardSidebar(
                
                # Dynamically creating the sidebar 
                sidebarMenuOutput("menu"),
                
                # Dynamically creating this section on sever side
                menuItemOutput("menuitem")
        ),
        
         # Create Body
        dashboardBody(
                
                tabItems(
                        
                        # First Sidebar Menu Tab content
                        tabItem(tabName = "dashboard",
                        
                        # First Row - "Dashboard"
                        fluidRow(
                                # Controlls for map
                                column(4, box(
                                               title = "controls", 
                                               sliderInput("num", 
                                                           label="Choose a number", 
                                                           value = 100, min = 10, max = 250),
                                               sliderInput("sd",
                                                           label = "Choose a sd", 
                                                           value = 10, min = 5, max = 15),
                                               sliderInput("var",
                                                           label = "Choose a variance",
                                                           value = 150, min = 100, max = 200))
                                       ),
                                
                                # render Map
                                column(8, box( collapsible = TRUE,
                                        title = "Labour Market Map",
                                        plotOutput("plot1",height = 350)
                                ))
                                
                        ),
                        # Second Row - "Dashboard"
                        fluidRow(
                                # This row will represent summary graphics for the chosen map layout
                                column(4),
                                       
                                
                                # Second graphic
                                column(4, 
                                       
                                       box(collapsible = TRUE,
                                           title = "plot2",
                                           plotOutput("plot2", height = 250, width = "100%"))
                                       ),
                                
                                # Third grapic
                                column(4,
                                       title = "controls",
                                       box( collapsible = TRUE,
                                            plotOutput("plot3", height = 250),
                                            p("This will display a map of Austria that i will put all information about the relevant players in a region into."))
                                       )
                                )
                        ),
                        
                
                
                
                # 2nd Sidebar-Menu Item - "Widgets"
                tabItem(tabName = "widgets", 
                        
                        # First Row of Content
                        fluidRow(
                                column(6, tabBox(
                                        tabPanel("Tab1", 
                                                 h4("Tab Header"),
                                                 p("Tab content. This can actually be filled with anything i would like to put in here.")
                                                 ),
                                        tabPanel("Tab2",
                                                 h4("Tab 2 Header"),
                                                 p("Again this can be set to hold any kind of information i would like to present here.")),
                                        tabPanel("Tab3", 
                                                 h4("Tab 3 Header"),
                                                 p("This too hold information for later use.s"))
                                ))
                        ),
                        
                        # Second Row of Content
                        fluidRow(
                                # Row1-Column1
                                column(6, box(
                                        title = "first Box",
                                        plotOutput("plot4", height = 200)
                                        
                                ))
                        )
                        ),
                
                # 3rd Sidebar-Menu Item
                tabItem(tabName = "data",
                        
                        # First Row of Content
                        fluidRow(
                                
                                # First column
                                column(5,
                                       tabBox(title = "Source Connections",
                                               tabPanel(
                                                       title = "Status",
                                                       h3("Current Datasource Status"),
                                                       p("This describes the current connection and retrieval status for all associated data sources.")
                                               ),
                                               tabPanel(
                                                       title = "Jobs",
                                                       h3("Current Job status on data retrieval"),
                                                       p("This describes the current progress and status on the data retrieval jobs scheduled."),
                                                       tabsetPanel(
                                                               tabPanel(
                                                                       title = "out of sync",
                                                                       h5("Datasources found to be updatable.")
                                                               ),
                                                               tabPanel(
                                                                       title = "up to date",
                                                                       h5("Datasoureces that are up to date.")
                                                               ),
                                                               tabPanel(
                                                                       title = "disconnected",
                                                                       h5("Datasources currently unavailable.")
                                                               )
                                                       )
                                               )
                                               
                               )),
                               # Second Column
                                column(7, 
                                       tabBox(title = "Data Retrieval",
                                              tabPanel(
                                                      title = "Trafic",
                                                      h3("Network Traffic"),
                                                          plotOutput(outputId = "datasourceTraffic",
                                                                     height = 450)
                                              ),
                                              tabPanel(
                                                      title = "Progress",
                                                      h3("Ongoing Traffic"),
                                                          plotOutput(outputId = "trafficOngoing",
                                                                     height = 250),
                                                      h3("Progress"),
                                                          plotOutput(outputId = "trafficProgress",
                                                                     height = 250)
                                              )
                                              ))
                        ),
                        # Second Row of Content
                        fluidRow(
                                # First Column
                                
                                # Second Column
                                # Third Column
                                
                )),
                
                # 4th Sidebar-Menu Item "Schedules"
                tabItem(tabName = "schedules",
                        
                        # First row of Content
                        fluidRow(
                                
                                box( title = ""
                                 
                                
                                )
                                
                        )
                        
                        ), #End of "Schedules"
                
                
                
                # 5th Sidebar-Menu Item 
                tabItem(tabName = "settings",
                        
                        # First row of Content
                        fluidRow(
                                
                                # First Column
                                column(4,
                                       box( title = "Color Setup",
                                       radioButtons(inputId = "color",
                                                          label = "Set the Dashboard Theme",
                                                          choices = c("black"= "black", 
                                                                      "white"="white", 
                                                                      "green"="green", 
                                                                      "red"="red", 
                                                                      "orange"="orange"))
                                       
                                       )
                                       )
                        ))
        
                )
       
        )
)

server <- function(input, output){
        
        # Dynamically creating content for the sidebar
        output$menu <- renderMenu({
            sidebarMenu(
                     sidebarSearchForm(textId = "searchText", buttonId = "searchButton",
                    label = "Search..."),
                    menuItem("Dashboard", tabName = "dashboard", icon = icon("dashboard")),
                    menuItem("Widgets", tabName = "widgets", icon = icon("th")),
                    menuItem("Data", tabName = "data", icon = icon("data")),
                    menuItem("Schedules", tabName = "schedules",icon = icon("calendar")),
                    menuItem("Settings", tabName = "settings", icon = icon("cog", lib = "glyphicon")),
                    
                    
                    # Include dynamic input items
                    sliderInput(inputId = "inp", label = "Slider:",
                                value = 10, min = 1, max = 20), 
                    
                    textInput(inputId = "text", label = "Text input:"),
                    
                    dateRangeInput(inputId = "daterng", label ="Date Range:", start = Sys.time())
            )
          })
        
}

shinyApp(ui, server)
```


