---
title: "Shiny in R Data Product Development"
output: html_notebook
---


```{r}
install.packages("shiny", dependencies=TRUE)
require(shiny)
```

### Basics in R Shiny 

#### How to build a Shiny App

- `ui.R` : Sets the user interface
- `server.R` controls what the app does 

Simple code example:
```{r}
shinyUI(fluidPage(
        titlePanel("Data science FTW!"),
        sidebarLayout(
                sidebarPanel(
                        h3("Sidebar Text")
                        ),
                mainPanel(
                        h3("Main Panel Text"),
                        p("Text here")
                        
                )
        )
))
```

## Most simple shiny code structure:
```{r}
#  Setting up the html structure
ui <- fluidPage(
        inputPanel()
)

# Setting up the server structure:
server <- function(input, output){}

# Combining in the shiny app
shinyApp(ui = ui, server= server)
```

Think about "inputs" and "outputs " when building a shiny app.

Input functions to use in shiny:

- `sliderInput()` : produces a slider input object to select a value from a given range

```{r}
ui <- fluidPage(
        titlePanel("This is the title!")
        
        sidebarLayout(
                sidebarPanel(
                        # Adding a slider input object
                        sliderInput(inputId = "num",
                                    label = "Choose a number",
                                    value = 25, 
                                    min = 1, max = 100)
                )
        )
        
        
        # radioButtons("radio", 
        #      "Take a choice", 
        #      choices = c("a"="a","b"= "b","c"= "c")
        #      )


        # Adding a group of checkbox objects 
        # checkboxGroupInput()
        
        # Adding a single Checkbox input object
        # checkboxInput()
        #  
        # dateInput()
        # dateRangeInput()
        # actionButton()
        # submitButton()
        # actionLink()
        # fileInput()
        # numericInput()
        # passwordInput()
        
     
        
        # textInput()
        
)


server <- function(input, output){}

shinyApp(ui, server)
```

- inputId = "num"
- label = "Description of the function"
- plus specific input arguments


## Output functions

- `dataTableOutput()`   : an interactive table
- `htmlOutput()`        : raw HTML
- `imageOutput()`       : image
- `plotOutput()`        : plot
- `tableOutput()`       : table  
- `textOutput()`        : text
- `uiOutput()`          : a shiny UI element
- `verbatimTextOutput()`: text

```{r}
ui <- fluidPage(
        
        # Create a title bar
        titlePanel("Title gets put here")
        
        # Create a Sidebar to hold the input functions
        
        
        
        # Creating a slider input function
        sliderInput(inputId = "num",
                    label = "Choose a number", 
                    value = 200, min =1, max = 10000),
        
        # Creating a slider to set the mean
        sliderInput(inputId = "mean",
                    label = "Set the expected mean",
                    value = 100, min = 20, max = 180),
        
        # Create a slider to set the sd 
        sliderInput(inputId = "sd",
                    label="Set the expected sd",
                    value = 20, min = 1, max = 80),
        
        # Creating a ouput function
        plotOutput("hist")
)

server <- function(input, output){
        
        # Define the histogram function to serve #hist
        output$hist <- renderPlot({ 
                title <- "100 random normal values"
                hist(rnorm(input$num, input$mean, input$sd), main = title) 
                
                })
}

shinyApp(ui, server)
```
Render Functions for creating the render object for the output:

- `renderDataTable()`   : An interactive Table
- `renderImage()`       : An image
- `renderPlot()`        : A plot
- `renderPrint()`       : A code block of printed output
- `renderTable()`       : a table
- `renderText()`        : A character string
- `renderUI()`          : a Shiny UI Element



## Deeper Look at `Reactivity`

Looking at **reactive values**
```{r}

# Setting up the UserInterface for the app
ui <- fluidPage(
        
        titlePanel(
                
                h3("Title here")
                ),
        
        # Creating a sidebar
        sidebarLayout(
                
                # Setting the sidebarPanel
                sidebarPanel(
                        h4("Sidebar Panel Title"),
                        
                        # Enter some description here
                        p("Some text here describing the app"),
                        
                        # Setting up the Input function
                        sliderInput(inputId = "num",
                                    label = "Set the n of observations",
                                    value = 20, min = 1, max= 200), 
                        
                        # Input for the expected mean value
                        sliderInput(inputId = "mean",
                                    label = "Set the mean",
                                    value = 100, min =1, max = 140),
                        
                        # Input for the expected sd in the population
                        sliderInput(inputId = "sd",
                                    label = "Set the sd",
                                    value = 15, min = 1, max = 50)
                        
                        
                        
                        
                ),
                
                
                        
                # Displaying the histogram plot
                # Setting the main Panel
                mainPanel(
                        
                        # Title
                        h2("Mainbar Title"),
                        
                        
                        # Displaying the output plot
                        plotOutput("hist")
                )
                
                
                
                
        )
        
        
)

server <- function(input, output){
        
        output$hist <- renderPlot({
                
                # Setting the title
                title <- "The gaussian standard normal distribution"
                
                hist(rnorm(input$num, input$mean, input$sd),
                     main = title)
        })
}

shinyApp(ui, server)
```

## The reactive toolkit in Shiny

Reactive Function use :

- a code chunk to build (and rebuild) an object
- this object will repsond to changes in a set of reactive values

#### `render` functions
```{r}
# Basic syntax: renderNAME { "function to be rendered" }

# rendering a plot
renderPlot({ hist(rnorm(input$num)) })

# rendering a data.table
renderDataTable()

# 
```

Code example
```{r}
ui <- fluidPage(
        
        sliderInput(
                inputId = "num", 
                label = "Choose a number", 
                value = 25, min = 1, max = 100
        ),
        
        textInput(
                inputId = "title",
                label = "Set title for the graph",
                value = "Histogram of Random Normal Values"),
        
        plotOutput("hist")
)

server <- function(input, output){
        
        output$hist <- renderPlot({
                hist(rnorm(input$num), main = input$title)
        })
}

shinyApp(ui, server)
```

### Modularize Code with `reactive()`


Instead of having two functions recreat their data based on a reactive input objects....
```{r}
ui <- fluidPage(
        
        sliderInput(
                inputId = "num",
                label = "Choose a number",
                value = 25, min = 1, max = 100
        ),
        
        # render a histogram
        plotOutput("hist"),
        
        # render a text block displaying the summary statistics
        verbatimTextOutput("stats")
)

server <- function(input, output){
        
        output$hist <- renderPlot({
                hist(rnorm(input$num))
        })
        
        output$stats <- renderPrint({
                summary(rnorm(input$num))
        })
}

shinyApp(ui, server)
```

... we would like for both render objects to display based on the same dataset. 
Instead both functions create a individual dataset based on the common link to the value of "num".
We can use `reactive()` to achieve this behaviour. 

Look out:

- you call a reactive expression like a function using `data()`
- Reactive expressions cache their values

```{r}
ui <- fluidPage(
        
        sliderInput(
                inputId = "num",
                label = "Choose a number",
                value = 25, min = 1, max = 100
        ),
        
        # render a histogram
        plotOutput("hist"),
        
        # render a text block displaying the summary statistics
        verbatimTextOutput("stats")
)

server <- function(input, output){
        
        # REACTIVE USED HERE !!
        rnorm_dat <- reactive( { rnorm(input$num) })
        
        output$hist <- renderPlot({
                hist(rnorm_dat())
        })
        
        output$stats <- renderPrint({
                summary(rnorm_dat())
        })
}

shinyApp(ui, server)
```
Still we get a lot of unnecessary updates when the user slides the input into possition or types a single letter into a textbox. 

We most often want to have a good amount of lag until we finally take the input as a change, to assert a recomputation of our function call.

This is done with `isolate()` like this:

isolate({ rnorm(input$num) })
```{r}
ui <- fluidPage(
        
        sliderInput(
                inputId = "num",
                label = "Choose a number",
                value = 25, min = 1, max = 100
        ),
        
        # get a text input to change the main plot title
        textInput(inputId = "title",
                  label = "Write a title",
                  value = "Histogram of Random Normal Values"),
        
        # render a histogram
        plotOutput("hist"),
        
        # render a text block displaying the summary statistics
        verbatimTextOutput("stats")
)

server <- function(input, output){
        
        rnorm_dat <- reactive( { rnorm(input$num) })
        
        # Compute function for #hist
        output$hist <- renderPlot({
                hist(rnorm_dat(), main = isolate(input$title ))
        })
        
        # Compute function for #stats 
        output$stats <- renderPrint({
                summary(rnorm_dat())
        })
}

shinyApp(ui, server)
```

## Trigger code execution : using `observeEvent()`

```{r}
require(shiny)

ui <- fluidPage(
        
        # Addding an action button 
        actionButton(
                inputId = "clicks",
                label = "Click me",
                value = "Clicked"
        )
)

server <- function(input, output){
        
        # Adding function to the button
        observeEvent(input$clicks, { print(as.numeric(input$clicks))
        })
}

shinyApp(ui, server)
```
Using event to trigger code rendering using : `observe()`
```{r}
# basic Syntax
observe({ print(input$clicks) })

```


## Delaying server reaction in Shiny: `eventReactive()`
Now we want to delay reacations in Shiny with `eventReactive()`


We now update our basic functio to update only once the "update" button is clicked actively by the user.
```{r}
ui <- fluidPage(
        
        sliderInput(inputId = "num",
                    label = "Choose a number",
                    value = 25, min = 1, max = 100),
        
        actionButton(inputId = "update",
                     label = "Update"),
        
        plotOutput("hist")
)

server <- function(input, output){
        
        # Calculating the data, once the "update" event has been triggered
        data <- eventReactive(input$update, {rnorm(input$num) })
        
        output$hist <- renderPlot({
                hist(rnorm(data()))
        })
}

shinyApp(ui, server)
```
Manage states with `reactive Values()`

This creates a list of reactive values to manipulate proigrammatically

```{r}
require(shiny)

ui <- fluidPage(
        actionButton(inputId = "norm", label = "Normal"),
        actionButton(inputId = "unif", label = "Uniform"),
        plotOutput("hist")
)

server <- function(input, output){
        rv <- reactiveValues(data = rnorm(100))
        
        observeEvent(input$norm, {rv$data <- rnorm(100) })
        observeEvent(input$unif, {rv$data <- runif(100) })
        
        output$hist <- renderPlot({
                hist(rv$data)
        })
}

shinyApp(ui, server)
```

